---
title: "data_01_exploration of Irish and French fishery data"
author: "WGCSE"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs, message=FALSE, echo=TRUE, warning=FALSE}
gc()
rm(list=ls())

library(RODBC)
library(tidyverse)
library(lattice)
library(lubridate)
library(mapplots)
library(sp)
library(sf)
library(reshape2)
library(readxl)
library(knitr)
library(DataCombine)
library(lattice)
library(icesTAF)
library(magrittr)
library(viridis)
library(sessioninfo)
library(stringr)
```
# Setup options

```{r setupoptions}
curr.year <- 2024
dat.year <- curr.year -1
fu.n <- "FU2021"

save.plots <- T # set it up as T or F whether you want to save the plots when running all the chunks or kniting the document
save.tables <- T # set it up as T or F whether you want to save the tables "propmature"" and the "L50" into .txt files when running all the chunks or knitting the document

# ewen plot
source("bootstrap/software/functions/required.funcs.r")
source("bootstrap/software/functions/ld1.plot.r")
```

# Introduction

The goals of this R Markdown document are to:

* Preprocess and explore FU2021 Intercatch data for France and Ireland Sampling programmes
* Write TAF data tables

Files before (inside "bootstrap/data/"):

* intercatch folder
  * france.lfd.wgcse2019.csv
  * fu2021.alfd.hist2024.csv
  * fu2021.alfd.series2024.csv
  * fu2021.df.all.csv
  * fu2021.nat.alfd.2024.csv
  * fu2021.sam4.2024.csv
  * int.land.csv
  * ireland.lfd.wgcse2022.csv
  * MSY_nep_stock.csv
  * NepLand_WGCSE2024.csv
  * qtr.int.land.csv

Files after (inside "data_01_exploration"):


* national_landings_FU2021.csv
* international_landings_FU2021.csv
* discard.data_FU2021.csv
* IRE_meansize_FU2021.csv
* exp.country_FU2021.csv
* prop.male_FU2021.csv

* FU2021_National Landings qtr.png
* FU2021_International Landings.png
* FU2021_Discard Rate.national.programme.png
* FU2021_IRE.commercial.lfd.png
* FU2021_Mean_weights_country.png
* FU2021_LFD_country.png
* FU2021_proportion_male_landings.png
* FU2021_LFD.plot.png


# Data Work Up
<a href="#top">Back to top</a>

This Rmarkdown document gives the exploitation pattern from the raised numbers from Irish Sampling programme.
Then load in the Frence fishery and sampling data up to 2018.
Then raises to the international landings for this stock.

## National Landings.
<a href="#top">Back to top</a>

```{r nat land, echo=TRUE, message=FALSE, fig.height=10, fig.width=7, warning=FALSE, fig.cap="FU2021 Irish  Landings (t) by quarter."}
nat <- read.csv("bootstrap/data/intercatch/NepLand_WGCSE2024.csv")

nat.land <- nat %>% filter(FU ==2021)

nat.land <- nat.land[c(1:8)]
nat.land$LiveWt <- nat.land$LiveWt/1000

nat.land$Quarter <- as.factor(nat.land$Quarter)

ggplot(nat.land, aes(Year, LiveWt))+ geom_col(aes(fill = Quarter)) + 
       theme_bw() + ylab("Landings in Tonnes") + 
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank()) +
  scale_x_continuous(breaks = c(seq(min(nat$Year), max(nat$Year), by = 5), max(nat$Year))) +
  scale_y_continuous(breaks = seq(0, 2500, 200))

if (save.plots == T) {
  ggsave("data/data_01_exploration/FU2021_national_landings_qtr.png")
}

nl <- nat.land %>% group_by(Year, Quarter) %>% summarise(tot =sum(LiveWt)) 

nl <- nl %>% spread( Quarter, -Year)

nl$Total<- rowSums(nl[,2:5], na.rm=TRUE)

knitr::kable(nl , digits=3)

fileOut <- paste("data/data_01_exploration/national_landings_", fu.n, ".csv", sep = "")

 if (save.tables == T) {
   write.table(nl, fileOut, sep=",", row.names = F)
 }

```


## International Landings 
<a href="#top">Back to top</a>

In 2019 downwards revision (-50%) to UK-England catch data.

```{r int land, echo=TRUE, fig.height=10, fig.width=7, message=FALSE, warning=FALSE, fig.cap="FU2021 International Landings (t) ."}
int.land <- read.csv("bootstrap/data/intercatch/int.land.csv")

knitr::kable(int.land , digits=4)

int.land <- gather(int.land,"country", "landings",2:5) 



pi <- ggplot(int.land, aes(Year, landings, fill=country)) + geom_bar(stat="identity", position= "stack") + 
  theme_bw() + ylab("Landings in Tonnes") +
  scale_fill_manual(breaks = c("France", "Belgium", "Ireland", "UK"),
                     values=c("#56B4E9","#E69F00","#009E73", "#CC6666")) +
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank()) +
  scale_x_continuous(breaks = c(seq(min(int.land$Year), max(int.land$Year), by = 5), max(int.land$Year))) +
  scale_y_continuous(breaks = seq(0, 3800, 200))

pi + theme(axis.text=element_text(size=10),axis.title=element_text(size=10)) 


if (save.plots == T) {
  ggsave("data/data_01_exploration/FU2021_International Landings.png")
}


int.land <- int.land %>% mutate_all(funs(replace(., is.na(.), 0)))

int <- int.land %>% group_by(Year) %>% summarise(total=sum(landings))
names(int) <- c("year","total")
int$year  <- as.numeric(int$year)

fileOut <- paste("data/data_01_exploration/international_landings_", fu.n, ".csv", sep = "")

 if (save.tables == T) {
   write.table(int, fileOut, sep=",", row.names = F)
 }

```


## Load Irish fishery data 
<a href="#top">Back to top</a>

Load in national raised annual LFD object for data series 2012-2021.
This is calculated in 03_FU2021_InternationalDataRaising.rmd file held on national network.
Aggregated data set.

```{r data load, echo=FALSE, results='true'}
#historic aggregated data SALY
alfd.hist <- read.csv("bootstrap/data/intercatch/fu2021.alfd.hist2024.csv")
alfd.hist <-  alfd.hist[ , c(-1)]


alfd.wg <- read.csv("bootstrap/data/intercatch/fu2021.alfd.series2024.csv")
alfd.wg <-  alfd.wg[ , c(-1)]
```


## Data Table of estimated discard rates by number and weight from the Irish national sampling programme.
<a href="#top">Back to top</a>

```{r dis data1, echo=TRUE, fig.height=7, fig.width=7, warning=FALSE}

hist1 <- alfd.hist[,c("year", "tot.ln", "tot.dn", "removals.n", "dead.discard.rate.n","discard.rate.w","discard.rate.n", "tot.lw", "tot.dw",  "mw.lan", "mw.dis")]


hist1 <- hist1[,c("year","tot.ln", "tot.dn", "removals.n", "discard.rate.n", "discard.rate.w", "tot.lw", "tot.dw", "mw.lan", "mw.dis")]  
names(hist1) <- c("year","lan.num","dis.num","removals.n", "dis.rate.n", "dis.rate.wgt","lan.wgt", "dis.wgt", "mw.lan", "mw.dis")



summ <- alfd.wg %>% group_by(year) %>% summarise(lan.num=sum(rlan.n)/1000,
                                                dis.num=sum(rdis.n)/1000,
                                                lan.wgt=sum(rlan.wt),
                                                 dis.wgt=sum(rdis.wt)) %>% 
                                      mutate(removals.n = dis.num *0.75 + lan.num,
                                             dis.rate.n=dis.num/(lan.num+dis.num)*100,
                                             dis.rate.wgt=dis.wgt/(lan.wgt+dis.wgt)*100,
                                              mw.lan = lan.wgt/lan.num,
                                              mw.dis = dis.wgt/dis.num)


summ <- summ[,c("year", "lan.num", "dis.num","removals.n","dis.rate.n", "dis.rate.wgt", "lan.wgt", "dis.wgt", "mw.lan", "mw.dis")]
summ1 <- rbind(as.data.frame(hist1),  as.data.frame(summ))

dis <- summ1

knitr::kable(dis , digits=4)

fileOut <- paste("data/data_01_exploration/discard.data_", fu.n, ".csv", sep = "")

 if (save.tables == T) {
   write.table(dis, fileOut, sep=",", row.names = F)
 }

```



## Plot the estimated discard rates by number and weight from the Irish national sampling programme.
<a href="#top">Back to top</a>

```{r dis data2, echo=TRUE, fig.height=7, fig.width=7, warning=FALSE, fig.cap="Discard rates by number (red-line) and weight (blue-line) from national sampling programme."}

viridis_colors <- viridis_pal()(3)

dis <- dis[c(1, 5:6)]
dis <- dis %>% gather("disw","var" ,2:3)
names(dis) <- c("year", "type", "rate")

p1 <- ggplot(dis, aes(year, rate, group = type, colour=factor(type))) + 
  geom_line(size = 1) + 
  coord_cartesian(ylim= c(0, 100)) +
      ylab("% Discard rate") + 
  scale_colour_viridis_d(name = "Type", option = "D", end = 0.9) +  # Using color-blind friendly palette
  theme_bw() +
  scale_x_continuous(breaks = c(seq(min(dis$year), max(dis$year), by = 2), max(dis$year))) +
  theme(axis.text=element_text(size=11),
        axis.title=element_text(size=11),
        legend.title=element_blank(),
        panel.grid=element_blank(),
        legend.position = "bottom")

if (save.plots == T) {
  ggsave("data/data_01_exploration/FU2021_Discard Rate.national.programme.png")
}

```


## Irish Length Frequency Distribution Plot.
<a href="#top">Back to top</a>

Plot the Irish commercial length frequency distribution.
This is from the "exel" workup as was presented at previous WGs.

```{r lfd, echo=FALSE, results='true', fig.height=10, fig.width=7, fig.cap="FU2021 Irish Raised LFD and mean size displayed where black line = mean size in landings and red line = mean size in discards."}
viridis_colors <- viridis_pal()(3)

rldf <- read.csv("bootstrap/data/intercatch/ireland.lfd.wgcse2022.csv") 

l <- alfd.wg %>% select("year","cl", "rlan.n", "sex") %>% mutate(type=c("landings")) %>% rename("n" = "rlan.n")
d <- alfd.wg %>% select("year","cl", "rdis.n", "sex") %>% mutate(type=c("discards")) %>% rename("n" = "rdis.n")
dat <- rbind(l,d)


rldf <- rbind(rldf, dat)

head(rldf)
#ggplot(rldf, aes(x=cl, y= n, fill=type)) + geom_bar(stat="identity") + facet_grid(year~sex)

p <- ggplot(rldf, aes(x = cl, y = n, fill = type)) + 
  geom_bar(stat = "identity") +
  scale_fill_viridis_d() 

rldf$cl.n <- rldf$cl * rldf$n

#get mean for lfd
vline.data <- rldf %>%
  group_by(year, sex, type) %>%
  summarise(mean = sum(cl.n) / sum(n))
#tail(vline.data)

fileOut <- paste("data/data_01_exploration/IRE.meansize_", fu.n, ".csv", sep = "")

 if (save.tables == T) {
   write.table(vline.data, fileOut, sep=",", row.names = F)
 }

vline.l <- filter(vline.data, type == "landings")
vline.d <- filter(vline.data, type == "discards")

p <- ggplot(rldf, aes(x = cl, y = n, fill = type)) + 
  geom_bar(stat = "identity") +
  scale_fill_manual(values = viridis_colors) +  # Using Viridis palette
  facet_grid(year ~ sex) + 
  geom_vline(aes(xintercept = mean), data = vline.l, colour = "black", linetype = "solid") +
  geom_vline(aes(xintercept = mean), data = vline.d, colour = "red", linetype = "solid") +
  theme_bw() + 
  theme(panel.grid = element_blank(), legend.position = "bottom")

p

if (save.plots == T) {
  ggsave("data/data_01_exploration/FU2021_IRE.commercial.lfd.png")
}

```

## Load French fishery data set.
<a href="#top">Back to top</a>

This is from the "exel" workup as was presented at previous WGs.
Read in French data provided by national expert via excel file for years (2012-2018).
No sample data from France since 2018.
Then present table the fishery summary for Ireland and France:
Landings and discards by number and weight, removals number, dead discard rate by number, discard rate by weight and number and finally 
mean weight in landings and discards.

```{r fra, echo=FALSE, results='true', fig.height=10, fig.width=7}

load("bootstrap/data/intercatch/france.fu2021summary.wgcse2019.rdata")

fr.dat <- c3
names(fr.dat) <- c("year","lan.num","dis.num","removals.n", "dead.discard.rate.n", "dis.rate.n", "dis.rate.wgt","lan.wgt", "dis.wgt", "mw.lan", "mw.dis")
fr.dat$nation <- "France"

summ2 <- summ1 %>% mutate(dead.discard.rate.n=(1- lan.num /removals.n)*100, .before=5) %>% mutate(nation="Ireland")


dat.all <- rbind(as.data.frame(fr.dat), as.data.frame(summ2))

#dead.discard.rate.n =  (1- lan.num /removals.n)*100,

knitr::kable(dat.all, digits=3)

fileOut <- paste("data/data_01_exploration/exp.country_", fu.n, ".csv", sep = "")

 if (save.tables == T) {
   write.table(dat.all, fileOut, sep=",", row.names = F)
 }

```



## Comparison of Irish and French mean weight data
<a href="#top">Back to top</a>

Plot the mean weights from both national fisheries.

```{r mw, echo=FALSE, results='true', fig.height=5, fig.width=7, fig.cap="Mean weight (gr) in the sampled catches by sex for each country."}
dat.wgt<- dat.all[ , c(1, 10:12)]


dat.wgt <- dat.wgt %>% gather("mw.lan", "var",2:3)
names(dat.wgt) <- c("year", "source", "type", "wgt")


pw <- ggplot(dat.wgt, aes(x=year, y=wgt, group=as.factor(type), colour=factor(type))) + 
        geom_line(size=.75)  +  theme_bw() + 
        ylab("Mean weight grs") +
        coord_cartesian(ylim=c(0, 85)) +
        facet_grid(~source) +
  scale_x_continuous(breaks = c(seq(min(dis$year), max(dis$year), by = 2), max(dis$year)))  +
  scale_colour_manual(values = viridis_colors)
 
pw + theme(axis.text=element_text(size=10),axis.title=element_text(size=12)) + 
  theme(legend.title=element_blank(), panel.grid=element_blank(), legend.position = "bottom") 
 

if (save.plots == T) {
  ggsave("data/data_01_exploration/FU2021_Mean_weights_country.png")
}

t <- dat.wgt %>% spread(source, wgt)

knitr::kable(t, digits=3)
```



## Commercial Length Frequency data for France and Ireland.
<a href="#top">Back to top</a>

Plot the mean weights from both national fisheries.
```{r comblfd, echo=FALSE, results='true', fig.height=10, fig.width=7, fig.cap="RLFD by country by type displaying mean size"}
fr.rldf <- read.csv("bootstrap/data/intercatch/france.lfd.wgcse2019.csv")
fr.rldf <-  fr.rldf[ ,c(-1)]
fr.rldf$source <- "France"

levels(fr.rldf$sex)

fr.rldf <- mutate_if(fr.rldf, is.factor, str_replace_all, pattern = "male", replacement = "Male")
fr.rldf <- mutate_if(fr.rldf, is.character, str_replace_all, pattern = "feMale", replacement = "Female")

names(fr.rldf) <- c("year",	"cl",	"n",	"sex", "type",	"source")

# Select relevant columns for Ireland and preprocess
ir.rldf <- rldf[, 1:5]
ir.rldf$source <- "Ireland"

# Combine datasets
dat <- rbind(ir.rldf, fr.rldf)

p <- ggplot(dat, aes(x=cl, y= n, fill=type)) + geom_bar(stat="identity") + facet_grid(year~source) + theme_bw()

p +  geom_vline(aes(xintercept= 25)) + geom_vline(aes(xintercept=35), linetype='dashed') + 
 theme(panel.grid=element_blank(), legend.position = "bottom") +
    scale_fill_viridis_d(name = "Type", option = "D", end = 0.9)

if (save.plots == T) {
  ggsave("data/data_01_exploration/FU2021_LFD_country.png")
}

```


## Proportion of Males in the landings.
<a href="#top">Back to top</a>

Report data table of proportion males in landings by number.
```{r pro, echo=FALSE, results='true', fig.height=5, fig.width=7, fig.cap="Proportion Males in the landings by country"}

pf <- dat %>% filter(type== 'landings', source=='France') %>% 
  group_by(year, sex) %>% 
  summarise(tot= sum(n)) %>% 
  spread(sex, tot)

pf$total<- rowSums(pf[,2:3], na.rm=TRUE)
names(pf) %<>% tolower
pf <- pf %>% 
      mutate(pro.m= round(male/total*100))
pf$country <- "France"


pi <- dat %>% 
  filter(type== 'landings', source=='Ireland') %>% 
  group_by(year, sex) %>% 
  summarise(tot= sum(n)) %>% 
  spread(sex, tot)

pi$total<- rowSums(pi[,2:3], na.rm=TRUE)
names(pi) %<>% tolower
pi<- pi %>% mutate(pro.m= round(male/total*100))
pi$country <- "Ireland"

pro <- rbind(pf, pi)

knitr::kable(pro, digits=3)

fileOut <- paste("data/data_01_exploration/prop.male_", fu.n, ".csv", sep = "")

 if (save.tables == T) {
   write.table(pro, fileOut, sep=",", row.names = F)
 }


p1 <- ggplot(pro, aes(as.factor(year), pro.m, group = country, colour=factor(country))) + geom_line(size = 1) +
  coord_cartesian(ylim= c(0, 100)) + 
  ylab("Proportion Male (%)") + xlab(" year") +
  theme_bw() +
  scale_colour_viridis_d(name = "Country", option = "D", end = 0.9)

p1 + theme(axis.text=element_text(size=10) +theme(legend.title=element_blank()) +
      theme(panel.grid=element_blank(), legend.position = "bottom"))
  

if (save.plots == T) {
  ggsave("data/data_01_exploration/FU2021_proportion_male_landings.png")
}

```

## Commercial Length Frequency Distributions
<a href="#top">Back to top</a>

Here we write the data  in the format for the ld1.plot.
The vertical lines indicate the minimum conservation reference size (25 mm) and the 35 mm visual reference level.

```{r ep , echo=TRUE, warning=FALSE, fig.height=8, fig.width=8, fig.cap="FU2021 Raised LFD and mean size displayed where black line = mean size in landings and red line = mean size in discards."}
lf <- dat

lf <- lf %>%
  mutate(sex = ifelse(sex == "male", "Male",
               ifelse(sex == "female", "Female", sex)))

lfg <- lf %>% group_by( year, cl, sex, type) %>% summarise(num = sum(n))
fu2021.ld <- spread(lfg, type, num)
fu2021.ld$Catch <- fu2021.ld$discards  + fu2021.ld$landings
names(fu2021.ld) <- c("Year",  "Length", "Sex", "Discards", "Landings", "Catch")
## here we write the CSV file in the fromat for the ld1.plot

##Ewen Plot
#get some length frequency data in
df <- as.data.frame(fu2021.ld)
#summary(df)

#the names of the data must be
#Year, Sex, Length, Landings, Discards, Catch
#in this instance they already are so we don't have to worry about it.
#call the function, parameter order is
#1 data frame
#2 title for the plot
#3 minimum year to start from
#4 maximum year to run to
#5 Minimum landing size
#6 reference line (typically 35mm so that we can see the proportion of the stock which is considered #fully selected.


png("data/data_01_exploration/FU2021_LFD.plot.png", height=1700, width=1200, res=200)

plot.ld(df, "FU2021", 2012, 2023, 25, 35)

dev.off()

plot.ld(df, "FU2021", 2012, 2023, 25, 35)

```


## Session
<a href="#top">Back to top</a>
```{r info , echo=TRUE, warning=FALSE, fig.height=8, fig.width=8}
session_info()

```
