---
title: "data_02_exploration of Irish effort data"
author: "WGCSE"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs, message=FALSE, echo=TRUE, warning=FALSE}
gc()
rm(list=ls())

library(RODBC)
library(tidyverse)
library(lattice)
library(lubridate)
library(mapplots)
library(sp)
library(sf)
library(reshape2)
library(readxl)
library(knitr)
library(DataCombine)
library(lattice)
library(icesTAF)
library(magrittr)
library(viridis)
library(sessioninfo)
library(stringr)

```


# Setup options

```{r setupoptions}
curr.year <- 2024
dat.year <- curr.year -1
fu.n <- "FU2021"

save.plots <- T # set it up as T or F whether you want to save the plots when running all the chunks or kniting the document
save.tables <- T # set it up as T or F whether you want to save the tables "propmature"" and the "L50" into .txt files when running all the chunks or knitting the document
```

# Introduction

The goals of this R Markdown document are to:

* Plot Irish effort data
* Write TAF data tables

Files before (inside "bootstrap/data"):


* effort folder
  * 2021_effort.ire.csv
  * Number of vessel by year and FU.csv
 

Files after (inside "data_02_irish_effortdata"):

* FU2021_effort_Kwdays.png
* FU2021_vessel_number.png


# Data Work Up
<a href="#top">Back to top</a>

This Rmarkdown document plots the effort data and number of vessels.


## Load Irish effort data and plot
<a href="#top">Back to top</a>

Load in  data series 2005-2022.Aggregated data set.
This is calculated and held on national network add next year data to the file.
WGCSE 2015 recommended that effort data in Kw days should be presented as these data are more informative than effort data uncorrected for vessel power.
Effort for Irish directed fleet (30% landings threshold).

```{r data eff, echo=TRUE, fig.height=7, fig.width=10, warning=FALSE, fig.cap="FU2021 Irish Effort (Kw days) by year."}
#historic aggregated data SALY

eff <- read.csv("bootstrap/data/effort/2021_effort.ire.csv")

eff <- eff %>% filter(year >= '2005')

neffplot <- ggplot(eff, aes(year, effort_kw_days)) + geom_line(size = 1) + geom_point() + theme_bw() +
             ylab("Effort in kW days ") + 
            coord_cartesian(ylim= c(0, 1200), xlim=c(2005,dat.year)) +
            scale_x_continuous(breaks = c(seq(min(eff$year), max(eff$year), by = 5), max(eff$year)))

neffplot + theme(axis.text=element_text(size=12),axis.title=element_text(size=12)) + 
  theme(legend.title=element_blank(), panel.grid=element_blank())


if (save.plots == T) {
  ggsave("data/data_02_irish_effortdata/FU2021_effort_Kwdays.png")
}

knitr::kable(eff , digits=3)
```



## Load Number of vessels reporting landings > 10 tonnes.
<a href="#top">Back to top</a>

```{r data ves, echo=TRUE, fig.height=7, fig.width=10, warning=FALSE, fig.cap="FU2021 Number of vessels with landings > 10 t."}

ves <- read.csv("bootstrap/data/effort/Number of vessel by year and FU.taf.csv")

ves <- ves %>% filter(FU =='2021', year >= '2005')

vesplot <- ggplot(ves, aes(year, Number_of_Vessels)) + geom_line(size = 1) + geom_point() + theme_bw() +
             ylab("Number of vessels with landings > 10 t ") + 
            coord_cartesian(ylim= c(0, 80), xlim=c(2005,dat.year)) +
            scale_x_continuous(breaks = c(seq(min(ves$year), max(ves$year), by = 5), max(ves$year)))

vesplot + theme(axis.text=element_text(size=12),axis.title=element_text(size=12)) + 
  theme(legend.title=element_blank(), panel.grid=element_blank())

if (save.plots == T) {
  ggsave("data/data_02_irish_effortdata/FU2021_number_vessels.png")
}

```


## Session
<a href="#top">Back to top</a>
```{r info , echo=TRUE, warning=FALSE, fig.height=8, fig.width=8}
session_info()

```


