---
title: "model_01_raising to international"
author: "WGCSE"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libs, message=FALSE, echo=TRUE, warning=FALSE}
gc()
rm(list=ls())

library(RODBC)
library(tidyverse)
library(lattice)
library(lubridate)
library(mapplots)
library(sp)
library(sf)
library(reshape2)
library(readxl)
library(knitr)
library(DataCombine)
library(lattice)
library(icesTAF)
library(magrittr)
library(viridis)
library(sessioninfo)
library(stringr)
```

# Setup options

```{r setupoptions}
curr.year <- 2024
dat.year <- curr.year -1
fu.n <- "FU2021"

save.plots <- T # set it up as T or F whether you want to save the plots when running all the chunks or kniting the document
save.tables <- T # set it up as T or F whether you want to save the tables "propmature"" and the "L50" into .txt files when running all the chunks or knitting the document
```

# Introduction

The goals of this R Markdown document are to:

* Raise national lfd to the international landings and calculate fishery summary table.
* Write TAF data tables


Files before (inside "data_01_exploration"):

* exp.country_FU2021.csv

* international_landings_FU2021.csv


Files after (inside "model_01"):

* fishery_summary_FU2021.csv
* fu2021_mean_wgt_source.png



## Load data files 
<a href="#top">Back to top</a>

National annual raised data file.
International landings data file
```{r load data, include=FALSE}

alfd.wg <- read.csv("data/data_01_exploration/exp.country_FU2021.csv")

int <- read.csv("data/data_01_exploration/international_landings_FU2021.csv")
```



## Now the International Raising Factor
<a href="#top">Back to top</a>

Scale the combined French and Irish fishery data to international landings.
Rsiing factors in the order of + 20%.
```{r raising int, echo=TRUE, warning=FALSE}

dat.comb <- alfd.wg %>% group_by(year) %>% summarise(tot.ln=sum(lan.num),
                                                 tot.dn=sum(dis.num),
                                                 tot.lw=sum(lan.wgt),                        
                                                 tot.dw=sum(dis.wgt))

                                                
int <- int %>% filter(year >='2012')
int.dat <- left_join(int, dat.comb, by="year")
int.dat$rf <- with(int.dat, total/tot.lw)

t <- int.dat[ , c("year", "rf")]

knitr::kable(t, digits=3)

```



## Assessment Summary data table
<a href="#top">Back to top</a>

Apply international raising factor and calculate fishery summary.
Landings and discards by weight (tonne) and number ('000s).
Estimated mean weight (gr) in landings and discards.
Discard rates by weight (tonne) and number ('000s) from the raised derived data using the annual discard ogive.
Removals in number.
Dead discard rate by weight (tonne) and number ('000s). 
Proportion Removals retained by number ('000s).
Discard survival estimated as 25%.
```{r summary table}

summ <- int.dat %>% mutate(int.lan.n = tot.ln*rf,
                                int.dis.n = tot.dn*rf,
                                int.lan.wt = total,
                                int.dis.wt = tot.dw*rf,
                                removals.n = (int.dis.n *0.75) + int.lan.n,
                                prop.removal.ret = int.lan.n /removals.n,
                                dead.discard.rate.n =  (1- int.lan.n /removals.n)*100,
                                discard.rate.n = int.dis.n/ (int.lan.n + int.dis.n )*100,
                              discard.rate.w = int.dis.wt/ (int.lan.wt + int.dis.wt )*100,
                              mw.lan =int.lan.wt/int.lan.n,
                              mw.dis=int.dis.wt/int.dis.n)

summ1 <- summ [ , c(1, 8:18)]

knitr::kable(summ1[, c(1:8)], digits=3)
knitr::kable(summ1[, c(1, 9:12)], digits=3)

fileOut <- paste("model/model_01_raising_to_international/fishery_summary_", fu.n, ".csv", sep = "")

if (save.tables == T) {
    write.table(summ1, fileOut, sep=",", row.names = F)
}
```


## Plot mean weights by source.

<a href="#top">Back to top</a>

Combined weight used in assessment.
```{r mw plot, echo=FALSE, results='true', fig.height=10, fig.width=7, fig.cap="Mean weight (gr) in the sampled catches by source"}

dat1 <- summ1[, c(1, 11, 12)]
dat1$nation <- "combined"
dat2 <- alfd.wg[, c(1, 10:12)]

dat3 <- rbind(dat1, dat2)


dat.wgt <-  dat3 %>% gather("mw.lan", "var",2:3)
names(dat.wgt) <- c("year", "source", "type", "wgt")

knitr::kable(dat.wgt, digits=3)

viridis_colors <- viridis_pal()(2)

p <- ggplot(dat.wgt, aes(x=year, y=wgt, group=as.factor(type), colour=factor(type))) + 
        geom_line(size=.75)  +  theme_bw() + 
        ylab("Mean weight (grs)") +
        coord_cartesian(ylim=c(0, 60)) +
        facet_wrap(~source) +
  scale_x_continuous(breaks = c(seq(min(dat.wgt$year), max(dat.wgt$year), by = 2), max(dat.wgt$year))) +
  scale_fill_viridis_d(name = "type", option = "D", end = 0.9)
 
p + theme(axis.text=element_text(size=9),axis.title=element_text(size=9)) + 
  theme(legend.title=element_blank(), panel.grid=element_blank(), legend.position = "bottom") 


if (save.plots == T) {
  ggsave("model/model_01_raising_to_international/fu2021_mean_wgt_source.png")
}

```



## Session
<a href="#top">Back to top</a>
```{r info , echo=TRUE, warning=FALSE, fig.height=8, fig.width=8}
session_info()
```
