---
title: "output_01_assessment_summary"
author: "WGCSE"
date: '`r Sys.Date()`'
output:
  html_document:
    toc: true
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libs, message=FALSE, echo=TRUE, warning=FALSE}
gc()
rm(list=ls())

library(RODBC)
library(tidyverse)
library(lattice)
library(lubridate)
library(mapplots)
library(sp)
library(sf)
library(reshape2)
library(readxl)
library(knitr)
library(DataCombine)
library(lattice)
library(icesTAF)
library(magrittr)
library(viridis)
library(sessioninfo)
library(stringr)
```

# Setup options

```{r setupoptions}
curr.year <- 2024
dat.year <- curr.year -1
fu.n <- "FU2021"

save.plots <- T # set it up as T or F whether you want to save the plots when running all the chunks or kniting the document
save.tables <- T # set it up as T or F whether you want to save the tables "propmature"" and the "L50" into .txt files when running all the chunks or knitting the document
```

# Introduction

The goals of this R Markdown document are to:

* Create Assessment summary Table and calculate Harvest Rate in the fishery year.
* Write TAF data tables

Files before (inside "model_01_raising_to_international"):

* fishery_summary_FU2021.csv


Files before (inside "model_02_kriging"):

* Labadie_Summary_ADG_method_2024.csv

Files after (inside "output_01_assessment_summary"):

* exp.wgcse_FU2021.csv
* nep.stock.wgmixfish_FU2021.csv
* FU2021 UWTV abundance estimates.png
* FU2021 Mean weights.png
* FU2021 Harvest_Rate.png



## Load data files 
<a href="#top">Back to top</a>

This UWTV data summary ADG file is calculated in Model_02 folder when new survey data is finalised and kriged to calculate abundance.
Updated UWTV survey series based on Rgeostats kriging method.
Abundance estimate is used to calculate harvest rate (removals.n/abundance) in the fishery.
Commercial fishery sampling programme data available since 2012.
UWTV survey series started in 2013. 

```{r uwtv table}

summ <-read.csv("model/model_01_raising_to_international/fishery_summary_FU2021.csv")


tv.res <- read.csv("model/model_02_kriging/Labadie_Summary_ADG_method_2024.csv")

tv <- tv.res[, c("Year", "abund", "ci", "upper", "lower")]
names(tv) <- c("year", "abund", "ci",  "upper", "lower")

# Create new row - Datacombine
New1 <- c(2012, NA, NA, NA, NA)
tv <- InsertRow(tv, NewRow = New1, RowNum = 1)

# calculate hr
exp <- right_join(summ, tv, by="year") %>% mutate(hr = removals.n/abund*100)
exp$ogive <- 'annual'


#put in same format for report/adg same as FU7- North Sea stocks
fu2021.exp <- exp[c("year", "int.lan.n", "int.dis.n", "int.lan.wt", "int.dis.wt", 
                    "removals.n", "prop.removal.ret", "dead.discard.rate.n", "discard.rate.n", "discard.rate.w", 
                    "mw.lan", "mw.dis", "abund", "ci", "upper", "lower", "hr" )]


knitr::kable(fu2021.exp[ ,c(1:8)] , digits=4)
knitr::kable(fu2021.exp[ ,c(1, 9:17)] , digits=4)


fileOut <- paste("output/output_01_assessment_summary/exp.wgcse_", fu.n, ".csv", sep = "")

if (save.tables == T) {
    write.table(fu2021.exp, fileOut, sep=",", row.names = F)
}
```

## Reformat for WGMIXFISH Stock object
<a href="#top">Back to top</a>

```{r mixfish table}

mix <- exp[c("year","abund", "ci", "int.lan.n", "int.dis.n", "removals.n", "hr", 
             "int.lan.wt", "int.dis.wt","discard.rate.n","dead.discard.rate.n",
             "mw.lan", "mw.dis","discard.rate.w", "prop.removal.ret")]

mix$survival.rate <- "0.25"
mix$survival.rate <- as.numeric(mix$survival.rate)

mix <- mix %>%
  add_column(fu = "fu.2021",
             .before = "year") %>%
  mutate(discard.rate.n = (discard.rate.n/100), 
         dead.discard.rate.n = (dead.discard.rate.n/100),
         discard.rate.w = (discard.rate.w/100))

names(mix) <- c("fu", "year", "abund", "ci", "landings.n", "discards.n", "removals.numbers", "harvest.rate", "landings.t", "discards.t", 
                "discard.rate.n", "dead.disc.rate.n", "mean.wt.lan.gr", "mean.wt.dis.gr", "discard.rate.wgt", "prop.removal.ret.n", "survival.rate")

fileOut <- paste("output/output_01_assessment_summary/nep.stock.wgmixfish_", fu.n, ".csv", sep = "")

if (save.tables == T) {
    write.table(mix, fileOut, sep=",", row.names = F)
}

tail(mix)
```


## Plot TV abundance and Confidance Intervals and MSY Btrigger.
<a href="#top">Back to top</a>

Btrigger reference point estimated at 450 million by WGCSE 2021.

```{r abund, echo=FALSE, warning=FALSE, fig.cap="FU2021 UWTV abundance estimates (millions of individuals) and Btrigger displayed as dashed line."}

tvd <- fu2021.exp %>% filter (year >= '2013')

    ggplot(tvd, aes(x=year, y= abund)) +
          theme_bw() +
          geom_errorbar(aes(ymax=upper, ymin=lower, width=0.25)) +
          geom_line(size = 1) +
          geom_point() +
          theme(panel.grid = element_blank()) +
          scale_x_continuous(name="\nYear",
                             breaks = seq(min(tvd$year), max(tvd$year), 1)) +
          scale_y_continuous(name = "Abundance (millions)\n",
                             breaks = seq(0, max(tvd$upper)+100, 250),
                             limits = c(0, max(tvd$upper)+100)) + 
          geom_hline(aes(yintercept=450),colour="#990000",linetype="dashed",size = 0.9) 

    
if (save.plots == T) {
  ggsave("output/output_01_assessment_summary/FU2021_UWTV_Survey.png")
}
```


## Plot of estimated mean weights from National sampling programme.
<a href="#top">Back to top</a>

Plot mean weights (grs) in landings and discards.

```{r mw, echo=FALSE, warning=FALSE, fig.cap="FU2021 Mean weights in grs series.Mean weight landings (black line) and mean weight discards (blue line)."}

mw <- fu2021.exp [ , c(1, 11,12)]
mw <- mw %>% gather("mw.lan", "var",2:3)
names(mw) <- c("year", "type", "wgt")

p3 <- ggplot(mw, aes(x=year, y=wgt, group=type, colour=factor(type))) + 
        geom_line(size=.75)  +  theme_bw() +
        xlab("\nYear") +
       scale_y_continuous(name = "Mean weight (gr)\n",
                          limits = c(0, 50),
                          breaks = seq(0, 50, 5)) +
        scale_colour_manual(labels=c("Discards mean weight", "Landings mean weight"), values=c("black","blue")) +
        theme(panel.grid=element_blank(), legend.position = "bottom") +
  scale_x_continuous(breaks = c(seq(min(mw$year), max(mw$year), by = 2), max(mw$year)))
  
 
p3 + theme(axis.text=element_text(size=10),axis.title=element_text(size=10)) + theme(legend.title=element_blank()) 

if (save.plots == T) {
  ggsave("output/output_01_assessment_summary/FU2021_Mean_weights.png")
}
```


## Plot of Harvest Rates Updated Survey series
<a href="#top">Back to top</a>

Current MSY Harvest Rate of 6.0 % for FU2021.

Refer to WGCSE FU2021 report section and recent Marine Institute UWTV report FU2021. 

```{r hr, echo=FALSE, warning=FALSE, fig.cap="FU2021 Harvest Rate series."}

p4 <- ggplot(fu2021.exp, aes(year, hr)) + geom_line(size = 1) +
    theme_bw() + coord_cartesian(ylim= c(0, 20), xlim=c(2012,dat.year)) + 
    ylab("Harvest rate percent") +
  scale_x_continuous(breaks = c(seq(min(fu2021.exp$year), max(fu2021.exp$year), by = 2), max(fu2021.exp$year)))


p4 + geom_hline(aes(yintercept=6.0),size = 1,colour="blue", linetype="dashed") + theme(axis.text=element_text(size=10),axis.title=element_text(size=10)) 

if (save.plots == T) {
  ggsave("output/output_01_assessment_summary/FU2021_Harvest_Rate.png")
}
```



## Session
<a href="#top">Back to top</a>
```{r info , echo=TRUE, warning=FALSE, fig.height=8, fig.width=8}
session_info()
```

